---
title: "p8105_hw6_qg2155"
author: "Qiu Xia (Helen) Guan"
date: "11/20/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
```

```{r page set up, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Question 1

### Importing dataset 
```{r import homocide}
url_base = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv" 

homocide_data_raw = read_csv(url_base)

homocide_data_raw

```
This raw dataset has `r nrow(homocide_data_raw)` rows and `r ncol(homocide_data_raw)` columns. The variables are uid, reported date, and victim data that includes last name, first name, race, age, and sex. Location-related variables are city, state, lat, lon. There is also a variable called disposition that describes the results of the case. For example, a homocide is a case that was closed by arrest whereas a case with sufficient evidence but an arrest was not possible (example: suspect died) is labeled as a closed case without arrest.

### Cleaning dataset 
* Create city_state variable
* Omit Dallas, TX; Phoenix, AZ; Kansas City, MO; Tulsa, AL
* Modify victim_race variable to be binary: White vs Non-White
* Create binary variable case_status: 
    * resolved: cases with disposition "Closed by arrest"
    * unresolved: cases with disposition "Closed without arrest" or "Open/No arrest"

```{r clean homocide}
homocide_data = homocide_data_raw %>% 
  mutate(victim_race = if_else(victim_race == "White", "White", "Non-white"),
         victim_race = fct_relevel(victim_race, "White"),
         victim_age = as.numeric(victim_age),
         case_status = if_else(disposition == "Closed by arrest", "Resolved", "Unresolved"),
         case_status = fct_relevel(case_status, "Resolved")) %>%
         mutate(city_state = paste(city,state, sep = ", ")) %>% 
         filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))
```

After omitting cities Dallas, Phoenix, Kansas City, and Tulsa, the cleaned dataset now has `r nrow(homocide_data)` observations. 

### Baltimore logistic regression

Logistic regression for odds ratio of solving homocides comparing non-white victims to white victims

```{r baltimore}
baltimore_data = homocide_data %>% 
  filter(city_state == "Baltimore, MD") 

baltimore_logistic = 
  baltimore_data %>% 
  glm(case_status ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         ci_l = exp(estimate - std.error * 1.96),
         ci_h = exp(estimate + std.error * 1.96)) %>% 
  select(term, log_OR = estimate, OR, ci_l, ci_h, p.value) %>% 
  filter(term == "victim_raceNon-white") %>% 
  knitr::kable(digits = 3)

baltimore_logistic
```

### Logistic regression for all cities

Odds ratio for solving homocides comparing non-white victims to white victims in every city

Create nested dataset using city_state

```{r nesting city_state}
city_nest =
  nest(homocide_data, uid:case_status) 
```

* The result is a list column for each of the `r nrow(city_nest)` cities.

Create function

```{r function city_state}
individual_city = function(data) {

  indi_city_glm = 
    data %>% 
    glm(case_status ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) %>% 
    broom::tidy() %>% 
    mutate(OR = exp(estimate),
           ci_l = exp(estimate - std.error * 1.96),
           ci_h = exp(estimate + std.error * 1.96)) %>% 
    select(term, log_OR = estimate, OR, ci_l, ci_h, p.value) %>% 
    filter(term == "victim_raceNon-white")
}
```

```{r city_state results}
city_results = 
  city_nest %>% 
  mutate(city_homocide = map(city_nest$data, individual_city)) %>% 
  unnest(city_homocide)

city_results
```

Create plot for OR of each city

```{r city plot}
city_results %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = ci_l, ymax = ci_h)) +
  labs(
    title = "Odds Ratio for solving homocides comparing non-white victims to white victims by city",
    x = "City",
    y = "Solving homocides OR"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Boston has the highest odds of solved homocides for non-white victims. 

## Question 2

### Importing dataset

```{r}
bw_data = read_csv(file = "./birthweight.csv") %>% 
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace)) 
```

This dataset has `r nrow(bw_data)` observations and `r ncol(bw_data)` variables. Categorical variables have been switched to factors. Using the commant skimr::skim, none of the variables seem to be missing any observations. 

```{r}
bw_data %>% 
  skimr::skim()
```

### Proposed linear model for birthweight (bwt)
mother's age at devlivery: momage, 
previous number of low birth weight babies: pnumlbw,
family monthly income: fincome,
gestational age in weeks: gaweeks

```{r}
bwt_linear = lm(bwt ~ momage + pnumlbw + fincome + gaweeks, data = bw_data) 
```

Graphing the proposed linear model

```{r}
bw_data %>% 
  add_residuals(bwt_linear) %>% 
  add_predictions(bwt_linear) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point()
```

Main effect and interaction models 

```{r}
bwt_maineff = lm(bwt ~ blength + gaweeks, data = bw_data)

bwt_interact = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = bw_data)
```

```{r}
cv_df = 
  crossv_mc(bw_data, 100) 
```

```{r}
cv_df = 
  cv_df %>% 
  mutate(bwt_linear = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         bwt_maineff = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         bwt_interact = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = .x))) %>% 
  mutate(rmse_linear = map2_dbl(bwt_linear, test, ~rmse(model = .x, data = .y)),
         rmse_maineff = map2_dbl(bwt_maineff, test, ~rmse(model = .x, data = .y)),
         rmse_interact = map2_dbl(bwt_interact, test, ~rmse(model = .x, data = .y)))
```

Plotting the rmse of each model

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>%  
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

The interaction model has the lowest rmse (287.4692) and therefore it is the best model. 

```{r}
rmse(bwt_linear, bw_data)
rmse(bwt_maineff, bw_data)
rmse(bwt_interact, bw_data)
```



